Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-06-28 manuscript/main.tex
proof of proposition 3
\label{proposition: eigenanalysis}

	The following proposition introduces the Tucker's Criterion adapted to our scalar-on-hybrid regression model setting based on the aformentioned operators defined in the hybrid space.
	\begin{proposition}
		\begin{equation*}
			\max \limits_{\substack{\boldsymbol{\xi} \in \mathcal{H} \\ \Vert \boldsymbol{\xi} \Vert_\mathcal{H}=1}} \textnormal{Cov}^2 \left(\langle \mathbf{W}, \boldsymbol{\xi} \rangle_\mathcal{H}, Y \right)
		\end{equation*}
		is achieved when $\boldsymbol{\xi}$ is an eigenfunction associated with the largest eigenvalue of $\mathcal{U}$.
		\label{proposition: eigenanalysis}
	\end{proposition}
	In other words, $l$-th PLS component $\boldsymbol{\xi}_l$ can be obtained as the \textit{first} eigenfunction of $\mathcal{U}^{[l]} = \Sigma_{YW}^{[l]} \otimes \Sigma_{YW}^{[l]}$ whose components are formulated using the $l$-th residualized response and predictors: $\Sigma_{YW}^{[l]} = E(Y^{[l]} \mathbf{W}^{[l]})$.

@@

	The following proposition states that the PLS components estimated from Proposition  \ref{Proposition: Regularized generalized rayleigh quotient} are orthonormal with respect to the modified inner product $\langle \langle \cdot \rangle \rangle$ that incorporates the roughness of the functions.
	\begin{proposition}
		The estimated PLS components, $\widehat{\boldsymbol{\xi}}_1, \widehat{\boldsymbol{\xi}}_2, \ldots, \widehat{\boldsymbol{\xi}}_L$, from the proposed PLS algorithm are mutually orthonormal with respect to the modified inner product $\langle \langle h_l, h_j \rangle \rangle = \langle h_l, h_j \rangle_{\mathcal{H}} + \sum_{k=1}^K \lambda_k \langle \ddot{f}_l^{(k)}, \ddot{f}_j^{(k)} \rangle_{L^2}$ in $\mathcal{H}$ that incorporates the roughness of the functions; i.e.,
		\begin{equation*}
			\langle \langle \widehat{\boldsymbol{\xi}}_l, \widehat{\boldsymbol{\xi}}_j \rangle \rangle 
			= \langle \widehat{\boldsymbol{\xi}}_l, \widehat{\boldsymbol{\xi}}_j \rangle_{\mathcal{H}} + \sum_{k=1}^K \lambda_k \left\langle \ddot{\widehat{\psi}}_l^{(k)}, \ddot{\widehat{\psi}}_j^{(k)} \right\rangle_{L^2}
			= \mathbb{1}(l=j). 
		\end{equation*}
		\label{proposition: modified orthnormality of PLS components}
	\end{proposition}

@@

	\vspace{1cm}
	\noindent $\boldsymbol{\cdot}$  \underline{\textbf{Proposition 3}}

	$\mathcal{U}$ is \textit{positive} because for every $\mathbf{h} \in \mathcal{H}$,
	\begin{equation*}
		\langle \mathcal{U} \mathbf{h}, \mathbf{h} \rangle_\mathcal{H} = \langle \langle \Sigma_{YW}, \mathbf{h} \rangle_{\mathcal{H}} \, \Sigma_{YW}, \mathbf{h} \rangle_\mathcal{H} = \langle \Sigma_{YW}, \mathbf{h} \rangle_\mathcal{H}^2 \ge 0.
	\end{equation*}

@@

	\vspace{1cm}
	\noindent $\boldsymbol{\cdot}$  \underline{\textbf{Proposition 4}}

	Before we prove Proposition 4, we state and prove the following lemma:
	\begin{lemma}
		$\mathcal{C}_{WY}$ is an adjoint operator of $\mathcal{C}_{YW}$. That is, $\mathcal{C}_{WY} = \mathcal{C}_{YW}^*$.
	\end{lemma}
	Proof.

@@

	\noindent Now we prove Proposition 4. The singular value decomposition of $\mathcal{C}_{YW}$ is given by
	\begin{equation*}
		\mathcal{C}_{YW} = \sum \limits_{j=1}^\infty \iota_j (f_{1j} \otimes f_{2j}).
	\end{equation*}
	Let $\Vert \cdot \Vert_{op}$ denote an operator norm. Then, applying Theorem 4.3.4 in \citet{Hsing2015}, we can show that
	\begin{equation*}
		\Vert \mathcal{C}_{YW} \Vert_{op} = \sup_{\substack{\mathbf{h} \in \mathcal{H} \\ \Vert \mathbf{h} \Vert_\mathcal{H}=1}} | \mathcal{C}_{YW} \mathbf{h} |^2 
		= \sup_{\substack{\mathbf{h} \in \mathcal{H} \\ \Vert \mathbf{h} \Vert_\mathcal{H}=1}} | E(\langle \mathbf{W}, \mathbf{h} \rangle_\mathcal{H} Y)|^2 
		= \sup_{\substack{\mathbf{h} \in \mathcal{H} \\ \Vert \mathbf{h} \Vert_\mathcal{H}=1}} \textnormal{Cov}^2(\langle \mathbf{W}, \mathbf{h} \rangle_\mathcal{H}, Y) 
		= \kappa_1^2,
	\end{equation*}
	with maximum attained at $\mathbf{h}=f_{11}$, which is an eigenfunction of  $\mathcal{C}_{YW}^* \, \circ \, \mathcal{C}_{YW} = \mathcal{C}_{WY} \, \circ \, \mathcal{C}_{YW} = \mathcal{U}$ corresponding to the largest eigenvalue $\kappa_1^2.$

@@

	Firstly, it is obvious by the unit-norm constraint enforced in Proposition 4 that:
	\begin{align*}
		\langle \langle \hat{\boldsymbol{\xi}}_l,  \hat{\boldsymbol{\xi}}_l \rangle \rangle &= \langle \widehat{\boldsymbol{\xi}}_l, \widehat{\boldsymbol{\xi}}_j \rangle_{\mathcal{H}} + \sum_{k=1}^K \lambda_k \left\langle \ddot{\widehat{\psi}}_l^{(k)}, \ddot{\widehat{\psi}}_j^{(k)} \right\rangle_{L^2}  \\
		&= \sum \limits_{k=1}^K \int_{\mathcal{T}_k} \widehat{\psi}_l^{(k)}(t_k) \widehat{\psi}_j^{(k)}(t_k) dt_k + \sum \limits_{r=1}^p \widehat{\theta}_{lr} \widehat{\theta}_{jr} + \sum_{k=1}^K \lambda_k \int_{\mathcal{T}_k} \ddot{\widehat{\psi}}_l^{(k)}(t_k) \ddot{\widehat{\psi}}_j^{(k)}(t_k) dt_k \\
		&= \sum \limits_{k=1}^K \int_{\mathcal{T}_k} \hat{\mathbf{a}}^{(k) \top}_l \mathbf{b}^{(k)}(t_k) \mathbf{b}^{(k) \top}(t_k) \hat{\mathbf{a}}^{(k)}_l  dt_k + \boldsymbol{\theta}_l^\top \boldsymbol{\theta}_l + \sum \limits_{k=1}^K \lambda_k \int_{\mathcal{T}_k} \hat{\mathbf{a}}^{(k) \top}_l \ddot{\mathbf{b}}^{(k)}(t_k) \ddot{\mathbf{b}}^{(k) \top}(t_k) \hat{\mathbf{a}}^{(k)}_l  dt_k \\
		& = \sum \limits_{k=1}^K \hat{\mathbf{a}}_l^{(k) \top} J^{(k)} \hat{\mathbf{a}}_l^{(k)} + \boldsymbol{\theta}_l^\top \boldsymbol{\theta}_l +  \sum \limits_{k=1}^K \lambda_k \hat{\mathbf{a}}^{(k) \top}_l \ddot{J}^{(k)} \hat{\mathbf{a}}^{(k)}_l \\
		& = \hat{\boldsymbol{\xi}}_l^{*\top} J^\ast \hat{\boldsymbol{\xi}}_l^* + \hat{\boldsymbol{\xi}}_l^{*\top} \Lambda \ddot{J}^\ast \hat{\boldsymbol{\xi}}_l^* = \hat{\boldsymbol{\xi}}_l^{*\top} (J^\ast +\Lambda \ddot{J}^\ast) \hat{\boldsymbol{\xi}}_l^* = 1
	\end{align*}
	where $\ddot{\mathbf{b}}^{(k)}(t_k) = (\ddot{b}_1^{(k)}(t_k), \ldots, \ddot{b}_M^{(k)}(t_k))^\top$

>>>>

	The following proposition introduces the Tucker's Criterion adapted to our scalar-on-hybrid regression model setting based on the aformentioned operators defined in the hybrid space.
	\begin{proposition}\label{proposition: eigenanalysis}
		\begin{equation*}
			\max \limits_{\substack{\boldsymbol{\xi} \in \mathcal{H} \\ \Vert \boldsymbol{\xi} \Vert_\mathcal{H}=1}} \textnormal{Cov}^2 \left(\langle \mathbf{W}, \boldsymbol{\xi} \rangle_\mathcal{H}, Y \right)
		\end{equation*}
		is achieved when $\boldsymbol{\xi}$ is an eigenfunction associated with the largest eigenvalue of $\mathcal{U}$.	
	\end{proposition}
	In other words, $l$-th PLS component $\boldsymbol{\xi}_l$ can be obtained as the \textit{first} eigenfunction of $\mathcal{U}^{[l]} = \Sigma_{YW}^{[l]} \otimes \Sigma_{YW}^{[l]}$ whose components are formulated using the $l$-th residualized response and predictors: $\Sigma_{YW}^{[l]} = E(Y^{[l]} \mathbf{W}^{[l]})$.

@@

	The following proposition states that the PLS components estimated from Proposition  \ref{Proposition: Regularized generalized rayleigh quotient} are orthonormal with respect to the modified inner product $\langle \langle \cdot \rangle \rangle$ that incorporates the roughness of the functions.
	\begin{proposition}\label{proposition: modified orthnormality of PLS components}
		The estimated PLS components, $\widehat{\boldsymbol{\xi}}_1, \widehat{\boldsymbol{\xi}}_2, \ldots, \widehat{\boldsymbol{\xi}}_L$, from the proposed PLS algorithm are mutually orthonormal with respect to the modified inner product $\langle \langle h_l, h_j \rangle \rangle = \langle h_l, h_j \rangle_{\mathcal{H}} + \sum_{k=1}^K \lambda_k \langle \ddot{f}_l^{(k)}, \ddot{f}_j^{(k)} \rangle_{L^2}$ in $\mathcal{H}$ that incorporates the roughness of the functions; i.e.,
		\begin{equation*}
			\langle \langle \widehat{\boldsymbol{\xi}}_l, \widehat{\boldsymbol{\xi}}_j \rangle \rangle 
			= \langle \widehat{\boldsymbol{\xi}}_l, \widehat{\boldsymbol{\xi}}_j \rangle_{\mathcal{H}} + \sum_{k=1}^K \lambda_k \left\langle \ddot{\widehat{\psi}}_l^{(k)}, \ddot{\widehat{\psi}}_j^{(k)} \right\rangle_{L^2}
			= \mathbb{1}(l=j). 
		\end{equation*}

@@

\section{Proof of Proposition \ref{Proposition: property of U}}\label{section:proof:Proposition: property of U}
	$\mathcal{U}$ is \textit{positive} because for every $\mathbf{h} \in \mathcal{H}$,
	\begin{equation*}
		\langle \mathcal{U} \mathbf{h}, \mathbf{h} \rangle_\mathcal{H} = \langle \langle \Sigma_{YW}, \mathbf{h} \rangle_{\mathcal{H}} \, \Sigma_{YW}, \mathbf{h} \rangle_\mathcal{H} = \langle \Sigma_{YW}, \mathbf{h} \rangle_\mathcal{H}^2 \ge 0.
	\end{equation*}

@@

	\section{Proof of Proposition \ref{proposition: eigenanalysis}}\label{section:proof:proposition: eigenanalysis}
	Before we prove Proposition \ref{proposition: eigenanalysis}, we state and prove the following lemma:
	\begin{lemma}
		$\mathcal{C}_{WY}$ is an adjoint operator of $\mathcal{C}_{YW}$. That is, $\mathcal{C}_{WY} = \mathcal{C}_{YW}^*$.
	\end{lemma}
	Proof.

@@

	\noindent Now we prove Proposition \ref{proposition: eigenanalysis}. The singular value decomposition of $\mathcal{C}_{YW}$ is given by
	\begin{equation*}
		\mathcal{C}_{YW} = \sum \limits_{j=1}^\infty \iota_j (f_{1j} \otimes f_{2j}).
	\end{equation*}
	Let $\Vert \cdot \Vert_{op}$ denote an operator norm. Then, applying Theorem 4.3.4 in \citet{Hsing2015}, we can show that
	\begin{equation*}
		\Vert \mathcal{C}_{YW} \Vert_{op} = \sup_{\substack{\mathbf{h} \in \mathcal{H} \\ \Vert \mathbf{h} \Vert_\mathcal{H}=1}} | \mathcal{C}_{YW} \mathbf{h} |^2 
		= \sup_{\substack{\mathbf{h} \in \mathcal{H} \\ \Vert \mathbf{h} \Vert_\mathcal{H}=1}} | E(\langle \mathbf{W}, \mathbf{h} \rangle_\mathcal{H} Y)|^2 
		= \sup_{\substack{\mathbf{h} \in \mathcal{H} \\ \Vert \mathbf{h} \Vert_\mathcal{H}=1}} \textnormal{Cov}^2(\langle \mathbf{W}, \mathbf{h} \rangle_\mathcal{H}, Y) 
		= \kappa_1^2,
	\end{equation*}
	with maximum attained at $\mathbf{h}=f_{11}$, which is an eigenfunction of  $\mathcal{C}_{YW}^* \, \circ \, \mathcal{C}_{YW} = \mathcal{C}_{WY} \, \circ \, \mathcal{C}_{YW} = \mathcal{U}$ corresponding to the largest eigenvalue $\kappa_1^2.$

@@

	Firstly, it is obvious by the unit-norm constraint enforced in Proposition \ref{proposition: eigenanalysis} that:
	\begin{align*}
		\langle \langle \hat{\boldsymbol{\xi}}_l,  \hat{\boldsymbol{\xi}}_l \rangle \rangle &= \langle \widehat{\boldsymbol{\xi}}_l, \widehat{\boldsymbol{\xi}}_j \rangle_{\mathcal{H}} + \sum_{k=1}^K \lambda_k \left\langle \ddot{\widehat{\psi}}_l^{(k)}, \ddot{\widehat{\psi}}_j^{(k)} \right\rangle_{L^2}  \\
		&= \sum \limits_{k=1}^K \int_{\mathcal{T}_k} \widehat{\psi}_l^{(k)}(t_k) \widehat{\psi}_j^{(k)}(t_k) dt_k + \sum \limits_{r=1}^p \widehat{\theta}_{lr} \widehat{\theta}_{jr} + \sum_{k=1}^K \lambda_k \int_{\mathcal{T}_k} \ddot{\widehat{\psi}}_l^{(k)}(t_k) \ddot{\widehat{\psi}}_j^{(k)}(t_k) dt_k \\
		&= \sum \limits_{k=1}^K \int_{\mathcal{T}_k} \hat{\mathbf{a}}^{(k) \top}_l \mathbf{b}^{(k)}(t_k) \mathbf{b}^{(k) \top}(t_k) \hat{\mathbf{a}}^{(k)}_l  dt_k + \boldsymbol{\theta}_l^\top \boldsymbol{\theta}_l + \sum \limits_{k=1}^K \lambda_k \int_{\mathcal{T}_k} \hat{\mathbf{a}}^{(k) \top}_l \ddot{\mathbf{b}}^{(k)}(t_k) \ddot{\mathbf{b}}^{(k) \top}(t_k) \hat{\mathbf{a}}^{(k)}_l  dt_k \\
		& = \sum \limits_{k=1}^K \hat{\mathbf{a}}_l^{(k) \top} J^{(k)} \hat{\mathbf{a}}_l^{(k)} + \boldsymbol{\theta}_l^\top \boldsymbol{\theta}_l +  \sum \limits_{k=1}^K \lambda_k \hat{\mathbf{a}}^{(k) \top}_l \ddot{J}^{(k)} \hat{\mathbf{a}}^{(k)}_l \\
		& = \hat{\boldsymbol{\xi}}_l^{*\top} J^\ast \hat{\boldsymbol{\xi}}_l^* + \hat{\boldsymbol{\xi}}_l^{*\top} \Lambda \ddot{J}^\ast \hat{\boldsymbol{\xi}}_l^* = \hat{\boldsymbol{\xi}}_l^{*\top} (J^\ast +\Lambda \ddot{J}^\ast) \hat{\boldsymbol{\xi}}_l^* = 1
	\end{align*}
	where $\ddot{\mathbf{b}}^{(k)}(t_k) = (\ddot{b}_1^{(k)}(t_k), \ldots, \ddot{b}_M^{(k)}(t_k))^\top$
